---
title: "PEGASUS Labs"
author: "MHuang"
date: "10/2/2020"
output: 
  html_document: 
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

## Packages

pkg <- c("tidyverse", "data.table", "here", "DT", "grid", "gridExtra")
for (p in pkg) {
        if (require(p, character.only = T)) {
                print(paste0(p, " loaded successfully"))
        } else {
                install.packages(p)
                require(p, character.only = T)
                print(paste0(p, " downloaded and loaded successfully"))
        }
}


## Paths

pegasus_dir <- "C:/Users/MHuang/Dropbox (Partners HealthCare)/PEGASUS/"
dat_dir <- "data/"
long_fmt_dir <- "data/long_format/"


## Filenames

rpdr_file_header <- str_c(pegasus_dir, "rpdr_data/rpdr_pull_102020_blood_diff_absolute/mq116_100920154626312764_")
rpdr_file_ending <- ".txt" 
all_subj_fname <- here(dat_dir, "all_subj_enc_bmi.csv")

```

# 1. Laboratory tests data
## 1.1. Read in data & parse out different labs

This is the hematology labs of blood differential absolute requested Oct 7, 2020

```{r}

all_subj <- fread(all_subj_fname)
all_subj[, Collect_Date := as.Date(Collect_Date, format = "%Y-%m-%d")]
all_subj[, .N, .(is.na(EMPI), is.na(Collect_Date))][order(is.na, is.na.1)] %>% 
        setnames(., c("missing_EMPI", "missing_Collect_Date", "N")) %>% print()

lab <- fread(str_c(rpdr_file_header, "Lab", rpdr_file_ending), quote = "")
lab[, unique(EMPI)] %in% all_subj[, EMPI] %>% table()
lab[, unique(EMPI)] %in% all_subj[SampleGroup == "Asthmatic", EMPI] %>% table()
lab[, Lab_Date := as.Date(substr(Seq_Date_Time, 1, 10), format = "%m/%d/%Y")]
lab[grepl(":", Seq_Date_Time), ':='(Lab_Hour = gsub("(.*) (.*):(.*)", "\\2", Seq_Date_Time), 
                                    Lab_Minute = gsub("(.*) (.*):(.*)", "\\3", Seq_Date_Time))]

lab <- merge(lab, all_subj[, .(EMPI, Collect_Date)], by = "EMPI", all.x = T)
lab[, lab_timegap_d := as.numeric(Collect_Date - Lab_Date)]
lab[, lab_timegap_y := lab_timegap_d/365.25]
lab[, summary(lab_timegap_y)]

lab_condensed <- unique(lab[, .(Group_Id, Loinc_Code, Test_Id, Test_Description, Specimen_Type)])

lab_condensed[grepl("(?i)eos", Group_Id), ] %>% datatable()
lab_condensed[grepl("(?i)eos", Group_Id), ][, .N, .(Group_Id, Loinc_Code)]
eos_code <- unique(lab_condensed[grepl("(?i)eos", Group_Id), Loinc_Code])

lab_condensed[grepl("(?i)neut", Group_Id), ] %>% datatable()
lab_condensed[grepl("(?i)neut", Group_Id), ][, .N, .(Group_Id, Loinc_Code)]
neut_code <- unique(lab_condensed[grepl("(?i)neut", Group_Id), ][!grepl("(?i)vacuo", Group_Id), Loinc_Code])
neut_vacuo_code <- unique(lab_condensed[grepl("(?i)neut", Group_Id), ][grepl("(?i)vacuo", Group_Id), Loinc_Code])

lab_condensed[grepl("(?i)baso", Group_Id), ] %>% datatable()
lab_condensed[grepl("(?i)baso", Group_Id), ][, .N, .(Group_Id, Loinc_Code)]
baso_code <- unique(lab_condensed[grepl("(?i)baso", Group_Id), Loinc_Code])

lab_condensed[grepl("(?i)lymph", Group_Id), ] %>% datatable()
lab_condensed[grepl("(?i)lymph", Group_Id), ][, .N, .(Group_Id, Loinc_Code)]
lymph_code <- unique(lab_condensed[grepl("(?i)lymph", Group_Id), ][!grepl("(?i)atypical", Group_Id), Loinc_Code])
lymph_atyp_code <- unique(lab_condensed[grepl("(?i)lymph", Group_Id), ][grepl("(?i)atypical", Group_Id), Loinc_Code])

lab_condensed[grepl("(?i)mono", Group_Id), ] %>% datatable()
lab_condensed[grepl("(?i)mono", Group_Id), ][, .N, .(Group_Id, Loinc_Code)]
mono_code <- unique(lab_condensed[grepl("(?i)mono", Group_Id), Loinc_Code])

lab_condensed[grepl("(?i)band", Group_Id), ] %>% datatable()
lab_condensed[grepl("(?i)band", Group_Id), ][, .N, .(Group_Id, Loinc_Code)]
band_code <- unique(lab_condensed[grepl("(?i)band", Group_Id), Loinc_Code])

lab_condensed[grepl("(?i)blast", Group_Id), ] %>% datatable()
lab_condensed[grepl("(?i)blast", Group_Id), ][, .N, .(Group_Id, Loinc_Code)]
blast_code <- unique(lab_condensed[grepl("(?i)blast", Group_Id), Loinc_Code])

lab_condensed[grepl("(?i)plasma", Group_Id), ] %>% datatable()
lab_condensed[grepl("(?i)plasma", Group_Id), ][, .N, .(Group_Id, Loinc_Code)]
plasma_cells_code <- unique(lab_condensed[grepl("(?i)plasma", Group_Id), Loinc_Code])

```

## 1.2. Subset out labs after serum collection

This will effectively exclude subjects with no serum collection date

```{r}

lab_trim <- lab[Lab_Date <= Collect_Date, ]
lab_trim[, length(unique(EMPI))] 
lab_trim[, unique(EMPI)] %in% all_subj[SampleGroup == "Asthmatic", EMPI] %>% table()

```

# 2. Eosinophil

```{r}

eos_long <- lab[Loinc_Code %in% eos_code, ]
eos_long[, unique(EMPI)] %in% all_subj[, EMPI] %>% table()


## Clean up eos counts data (test results & unit, missingness etc.)

eos_long[, ':='(eos_timegap_d = as.numeric(Collect_Date - Lab_Date), 
                eos = as.numeric(Result), 
                eos_date = Lab_Date)]
eos_long[is.na(eos), .N, Result] %>% datatable()
eos_long[grep("/", Result), ':='(eos = gsub("([0-9]+)/(.*)", "\\1", Result) %>% as.numeric(), 
                                 Reference_Units = gsub("([0-9]+)/(.*)", "/\\2", Result))]

eos_long[, .N, Reference_Units] %>% datatable()
eos_long[Reference_Units == "", .N, .(Result, eos)] %>% datatable()
eos_long[!is.na(eos), .N, .(eos < 5, Reference_Units)][order(eos, Reference_Units)] %>% datatable()
# eos_long[eos >= 5, ] %>% datatable()
# eos_long[eos >= 5 & grepl("^[K|th]", Reference_Units) & Abnormal_Flag == "", 
#          .(EMPI, Group_Id, Test_Description, Result, Abnormal_Flag, Reference_Units, Reference_Range, eos)]
eos_long[eos >= 5 & grepl("^[K|th]", Reference_Units) & Abnormal_Flag == "", 
         eos := eos/1000]
eos_long[eos >= 5 & Reference_Units %in% c("/CMM", "/CUMM", "/cumm", "UL"), eos := eos/1000]
eos_long[!is.na(eos), ':='(eos_unit = "K/uL", 
                           eos_ref_range = Reference_Range, 
                           eos_abnormal_flg = Abnormal_Flag)]

eos_long <- eos_long[!is.na(eos) & !is.na(Collect_Date), ]

```

## 2.1. Extract last record before serum collection

```{r}

eos_long[, summary(eos_timegap_d)]
eos_long[, ':='(eos_timegap_y = eos_timegap_d/365.25, 
                eos_timegap_d_abs = abs(eos_timegap_d))]
eos_long[, summary(eos_timegap_y)]

setkey(eos_long, EMPI, eos_timegap_d)
eos_lastbf <- eos_long[eos_timegap_d >= 0, ][, head(.SD, 1), EMPI]
eos_lastbf[, .N, .(eos_timegap_d <= 7, eos_timegap_y <= 1)] %>% 
        setnames(., c("eos_timegap<=7d_before", "eos_timegap<=1y_before", "N")) %>%
        print()
ggplot(eos_lastbf) + 
        geom_histogram(aes(eos_timegap_y), bins = 200) + 
        labs(title = "Timing of eosinophil record before/on serum collection", 
             # subtitle = "Time gap = 0 removed", 
             x = "Time gap in years", 
             caption = "n = 951") + 
        theme_minimal() + 
        theme(axis.title = element_text(size = 12), 
              axis.text = element_text(size = 12), 
              plot.caption = element_text(size = 12))
ggplot(eos_lastbf) +
        geom_histogram(aes(eos), bins = 200) +
        labs(title = "Eosinophil count last before/on serum collection",
             x = "Eos (K/uL)",
             caption = "n = 951") +
        theme_minimal() +
        theme(axis.title = element_text(size = 12),
              axis.text = element_text(size = 12),
              plot.caption = element_text(size = 12))

```

## 2.2. Extract record closest to serum collection

```{r}

setkey(eos_long, EMPI, eos_timegap_d_abs)
eos_closest <- eos_long[, head(.SD, 1), EMPI]
eos_closest[, .N, .(eos_timegap_d_abs <= 7, abs(eos_timegap_y) <= 1)] %>% 
        setnames(., c("eos_absolute_timegap<=7d", "eos_absolute_timegap<=1y", "N")) %>%
        print()
ggplot(eos_closest) + 
        geom_histogram(aes(eos_timegap_y), bins = 200) + 
        labs(title = "Timing of eosinophil record closest to serum collection", 
             # subtitle = "Time gap = 0 removed", 
             x = "Time gap in years", 
             caption = "n = 1045") + 
        theme_minimal() + 
        theme(axis.title = element_text(size = 12), 
              axis.text = element_text(size = 12), 
              plot.caption = element_text(size = 12))
ggplot(eos_closest) +
        geom_histogram(aes(eos), bins = 200) +
        labs(title = "Eosinophil count closest to serum collection",
             x = "Eos (K/uL)",
             caption = "n = 1045") +
        theme_minimal() +
        theme(axis.title = element_text(size = 12),
              axis.text = element_text(size = 12),
              plot.caption = element_text(size = 12))

```

# 3. Neutrophil

```{r}

neut_long <- lab[Loinc_Code %in% neut_code, ]
neut_long[, unique(EMPI)] %in% all_subj[, EMPI] %>% table()


## Clean up neut counts data (test results & unit, missingness etc.)

neut_long[, ':='(neut_timegap_d = as.numeric(Collect_Date - Lab_Date), 
                 neut = as.numeric(Result), 
                 neut_date = Lab_Date)]
neut_long[is.na(neut), .N, Result] %>% datatable()

neut_long[, .N, Reference_Units] %>% datatable()
neut_long[Reference_Units %in% c("", "UL"), .N, .(Result, neut, Reference_Units, Reference_Range, 
                                                  Abnormal_Flag)] %>% datatable()
neut_long <- neut_long[!(Reference_Units %in% c("%", "UL")), ]
neut_long[!is.na(neut), ':='(neut_unit = "K/uL", 
                             neut_ref_range = Reference_Range, 
                             neut_abnormal_flg = Abnormal_Flag)]

neut_long <- neut_long[!is.na(neut) & !is.na(Collect_Date), ]

```

## 3.1. Extract last record before serum collection

```{r}

neut_long[, summary(neut_timegap_d)]
neut_long[, ':='(neut_timegap_y = neut_timegap_d/365.25, 
                 neut_timegap_d_abs = abs(neut_timegap_d))]
neut_long[, summary(neut_timegap_y)]

setkey(neut_long, EMPI, neut_timegap_d)
neut_lastbf <- neut_long[neut_timegap_d >= 0, ][, head(.SD, 1), EMPI]
neut_lastbf[, .N, .(neut_timegap_d <= 7, neut_timegap_y <= 1)] %>% 
        setnames(., c("neut_timegap<=7d_before", "neut_timegap<=1y_before", "N")) %>%
        print()
ggplot(neut_lastbf) + 
        geom_histogram(aes(neut_timegap_y), bins = 200) + 
        labs(title = "Timing of neutrophil record before/on serum collection", 
             # subtitle = "Time gap = 0 removed", 
             x = "Time gap in years", 
             caption = "n = 951") + 
        theme_minimal() + 
        theme(axis.title = element_text(size = 12), 
              axis.text = element_text(size = 12), 
              plot.caption = element_text(size = 12))
ggplot(neut_lastbf) +
        geom_histogram(aes(neut), bins = 200) +
        labs(title = "Neutrophil count last before/on serum collection",
             x = "Neut (K/uL)",
             caption = "n = 951") +
        theme_minimal() +
        theme(axis.title = element_text(size = 12),
              axis.text = element_text(size = 12),
              plot.caption = element_text(size = 12))

```

## 3.2. Extract record closest to serum collection

```{r}

setkey(neut_long, EMPI, neut_timegap_d_abs)
neut_closest <- neut_long[, head(.SD, 1), EMPI]
neut_closest[, .N, .(neut_timegap_d_abs <= 7, abs(neut_timegap_y) <= 1)] %>% 
        setnames(., c("neut_absolute_timegap<=7d", "neut_absolute_timegap<=1y", "N")) %>%
        print()
ggplot(neut_closest) + 
        geom_histogram(aes(neut_timegap_y), bins = 200) + 
        labs(title = "Timing of neutrophil record closest to serum collection", 
             # subtitle = "Time gap = 0 removed", 
             x = "Time gap in years", 
             caption = "n = 1052") + 
        theme_minimal() + 
        theme(axis.title = element_text(size = 12), 
              axis.text = element_text(size = 12), 
              plot.caption = element_text(size = 12))
ggplot(neut_closest) +
        geom_histogram(aes(neut), bins = 200) +
        labs(title = "Neutrophil count closest to serum collection",
             x = "Neut (K/uL)",
             caption = "n = 1052") +
        theme_minimal() +
        theme(axis.title = element_text(size = 12),
              axis.text = element_text(size = 12),
              plot.caption = element_text(size = 12))

```

# 4. Basophil

```{r}

baso_long <- lab[Loinc_Code %in% baso_code, ]
baso_long[, unique(EMPI)] %in% all_subj[, EMPI] %>% table()


## Clean up baso counts data (test results & unit, missingness etc.)

baso_long[, ':='(baso_timegap_d = as.numeric(Collect_Date - Lab_Date), 
                 baso = as.numeric(Result), 
                 baso_date = Lab_Date)]
baso_long[is.na(baso), .N, Result] %>% datatable()

baso_long[, .N, Reference_Units] %>% datatable()
baso_long[Reference_Units == "", .N, .(Result, baso, Reference_Units, Reference_Range, 
                                       Abnormal_Flag)] %>% datatable()
baso_long[!is.na(baso), ':='(baso_unit = "K/uL", 
                             baso_ref_range = Reference_Range, 
                             baso_abnormal_flg = Abnormal_Flag)]

baso_long <- baso_long[!is.na(baso) & !is.na(Collect_Date), ]

```

## 4.1. Extract last record before serum collection

```{r}

baso_long[, summary(baso_timegap_d)]
baso_long[, ':='(baso_timegap_y = baso_timegap_d/365.25, 
                 baso_timegap_d_abs = abs(baso_timegap_d))]
baso_long[, summary(baso_timegap_y)]

setkey(baso_long, EMPI, baso_timegap_d)
baso_lastbf <- baso_long[baso_timegap_d >= 0, ][, head(.SD, 1), EMPI]
baso_lastbf[, .N, .(baso_timegap_d <= 7, baso_timegap_y <= 1)] %>% 
        setnames(., c("baso_timegap<=7d_before", "baso_timegap<=1y_before", "N")) %>%
        print()
ggplot(baso_lastbf) + 
        geom_histogram(aes(baso_timegap_y), bins = 200) + 
        labs(title = "Timing of basophil record before/on serum collection", 
             # subtitle = "Time gap = 0 removed", 
             x = "Time gap in years", 
             caption = "n = 951") + 
        theme_minimal() + 
        theme(axis.title = element_text(size = 12), 
              axis.text = element_text(size = 12), 
              plot.caption = element_text(size = 12))
ggplot(baso_lastbf) +
        geom_histogram(aes(baso), bins = 200) +
        labs(title = "Basophil count last before/on serum collection",
             x = "Baso (K/uL)",
             caption = "n = 951") +
        theme_minimal() +
        theme(axis.title = element_text(size = 12),
              axis.text = element_text(size = 12),
              plot.caption = element_text(size = 12))

```

## 4.2. Extract record closest to serum collection

```{r}

setkey(baso_long, EMPI, baso_timegap_d_abs)
baso_closest <- baso_long[, head(.SD, 1), EMPI]
baso_closest[, .N, .(baso_timegap_d_abs <= 7, abs(baso_timegap_y) <= 1)] %>% 
        setnames(., c("baso_absolute_timegap<=7d", "baso_absolute_timegap<=1y", "N")) %>%
        print()
ggplot(baso_closest) + 
        geom_histogram(aes(baso_timegap_y), bins = 200) + 
        labs(title = "Timing of basophil record closest to serum collection", 
             # subtitle = "Time gap = 0 removed", 
             x = "Time gap in years", 
             caption = "n = 1046") + 
        theme_minimal() + 
        theme(axis.title = element_text(size = 12), 
              axis.text = element_text(size = 12), 
              plot.caption = element_text(size = 12))
ggplot(baso_closest) +
        geom_histogram(aes(baso), bins = 200) +
        labs(title = "Basophil count closest to serum collection",
             x = "Baso (K/uL)",
             caption = "n = 1046") +
        theme_minimal() +
        theme(axis.title = element_text(size = 12),
              axis.text = element_text(size = 12),
              plot.caption = element_text(size = 12))

```

# 5. Lymphocyte

```{r}

lymph_long <- lab[Loinc_Code %in% lymph_code, ]
lymph_long[, unique(EMPI)] %in% all_subj[, EMPI] %>% table()


## Clean up lymph counts data (test results & unit, missingness etc.)

lymph_long[, ':='(lymph_timegap_d = as.numeric(Collect_Date - Lab_Date), 
                 lymph = as.numeric(Result), 
                 lymph_date = Lab_Date)]
lymph_long[is.na(lymph), .N, Result] %>% datatable()

lymph_long[, .N, Reference_Units] %>% datatable()
lymph_long[Reference_Units == "", .N, .(Result, lymph, Reference_Units, Reference_Range, 
                                        Abnormal_Flag)] %>% datatable()
lymph_long[Reference_Units %in% c("#/mm3", "/uL", "/mm3"), .N, 
           .(Result, lymph, Reference_Units, Reference_Range, Abnormal_Flag)] %>% datatable()
lymph_long[Reference_Units %in% c("#/mm3", "/uL", "/mm3"), lymph := lymph/1000]

lymph_long[!is.na(lymph), ':='(lymph_unit = "K/uL", 
                               lymph_ref_range = Reference_Range, 
                               lymph_abnormal_flg = Abnormal_Flag)]

lymph_long <- lymph_long[!is.na(lymph) & !is.na(Collect_Date), ]

```

## 5.1. Extract last record before serum collection

```{r}

lymph_long[, summary(lymph_timegap_d)]
lymph_long[, ':='(lymph_timegap_y = lymph_timegap_d/365.25, 
                  lymph_timegap_d_abs = abs(lymph_timegap_d))]
lymph_long[, summary(lymph_timegap_y)]

setkey(lymph_long, EMPI, lymph_timegap_d)
lymph_lastbf <- lymph_long[lymph_timegap_d >= 0, ][, head(.SD, 1), EMPI]
lymph_lastbf[, .N, .(lymph_timegap_d <= 7, lymph_timegap_y <= 1)] %>% 
        setnames(., c("lymph_timegap<=7d_before", "lymph_timegap<=1y_before", "N")) %>%
        print()
ggplot(lymph_lastbf) + 
        geom_histogram(aes(lymph_timegap_y), bins = 200) + 
        labs(title = "Timing of lymphocyte record before/on serum collection", 
             # subtitle = "Time gap = 0 removed", 
             x = "Time gap in years", 
             caption = "n = 960") + 
        theme_minimal() + 
        theme(axis.title = element_text(size = 12), 
              axis.text = element_text(size = 12), 
              plot.caption = element_text(size = 12))
ggplot(lymph_lastbf) +
        geom_histogram(aes(lymph), bins = 200) +
        labs(title = "Lymphocyte count last before/on serum collection",
             x = "Lymph (K/uL)",
             caption = "n = 960") +
        theme_minimal() +
        theme(axis.title = element_text(size = 12),
              axis.text = element_text(size = 12),
              plot.caption = element_text(size = 12))

```

## 5.2. Extract record closest to serum collection

```{r}

setkey(lymph_long, EMPI, lymph_timegap_d_abs)
lymph_closest <- lymph_long[, head(.SD, 1), EMPI]
lymph_closest[, .N, .(lymph_timegap_d_abs <= 7, abs(lymph_timegap_y) <= 1)] %>% 
        setnames(., c("lymph_absolute_timegap<=7d", "lymph_absolute_timegap<=1y", "N")) %>%
        print()
ggplot(lymph_closest) + 
        geom_histogram(aes(lymph_timegap_y), bins = 200) + 
        labs(title = "Timing of lymphocyte record closest to serum collection", 
             # subtitle = "Time gap = 0 removed", 
             x = "Time gap in years", 
             caption = "n = 1052") + 
        theme_minimal() + 
        theme(axis.title = element_text(size = 12), 
              axis.text = element_text(size = 12), 
              plot.caption = element_text(size = 12))
ggplot(lymph_closest) +
        geom_histogram(aes(lymph), bins = 200) +
        labs(title = "Lymphocyte count closest to serum collection",
             x = "Lymph (K/uL)",
             caption = "n = 1052") +
        theme_minimal() +
        theme(axis.title = element_text(size = 12),
              axis.text = element_text(size = 12),
              plot.caption = element_text(size = 12))

```

# 6. Monocyte

```{r}

mono_long <- lab[Loinc_Code %in% mono_code, ]
mono_long[, unique(EMPI)] %in% all_subj[, EMPI] %>% table()


## Clean up mono counts data (test results & unit, missingness etc.)

mono_long[, ':='(mono_timegap_d = as.numeric(Collect_Date - Lab_Date), 
                 mono = as.numeric(Result), 
                 mono_date = Lab_Date)]
mono_long[is.na(mono), .N, Result] %>% datatable()

mono_long[, .N, Reference_Units] %>% datatable()
mono_long[Reference_Units == "", .N, .(Result, mono, Reference_Units, Reference_Range, 
                                       Abnormal_Flag)] %>% datatable()

mono_long[!is.na(mono), ':='(mono_unit = "K/uL", 
                             mono_ref_range = Reference_Range, 
                             mono_abnormal_flg = Abnormal_Flag)]

mono_long <- mono_long[!is.na(mono) & !is.na(Collect_Date), ]

```

## 6.1. Extract last record before serum collection

```{r}

mono_long[, summary(mono_timegap_d)]
mono_long[, ':='(mono_timegap_y = mono_timegap_d/365.25, 
                 mono_timegap_d_abs = abs(mono_timegap_d))]
mono_long[, summary(mono_timegap_y)]

setkey(mono_long, EMPI, mono_timegap_d)
mono_lastbf <- mono_long[mono_timegap_d >= 0, ][, head(.SD, 1), EMPI]
mono_lastbf[, .N, .(mono_timegap_d <= 7, mono_timegap_y <= 1)] %>% 
        setnames(., c("mono_timegap<=7d_before", "mono_timegap<=1y_before", "N")) %>%
        print()
ggplot(mono_lastbf) + 
        geom_histogram(aes(mono_timegap_y), bins = 200) + 
        labs(title = "Timing of monocyte record before/on serum collection", 
             # subtitle = "Time gap = 0 removed", 
             x = "Time gap in years", 
             caption = "n = 954") + 
        theme_minimal() + 
        theme(axis.title = element_text(size = 12), 
              axis.text = element_text(size = 12), 
              plot.caption = element_text(size = 12))
ggplot(mono_lastbf) +
        geom_histogram(aes(mono), bins = 200) +
        labs(title = "Monocyte count last before/on serum collection",
             x = "Mono (K/uL)",
             caption = "n = 954") +
        theme_minimal() +
        theme(axis.title = element_text(size = 12),
              axis.text = element_text(size = 12),
              plot.caption = element_text(size = 12))

```

## 6.2. Extract record closest to serum collection

```{r}

setkey(mono_long, EMPI, mono_timegap_d_abs)
mono_closest <- mono_long[, head(.SD, 1), EMPI]
mono_closest[, .N, .(mono_timegap_d_abs <= 7, abs(mono_timegap_y) <= 1)] %>% 
        setnames(., c("mono_absolute_timegap<=7d", "mono_absolute_timegap<=1y", "N")) %>%
        print()
ggplot(mono_closest) + 
        geom_histogram(aes(mono_timegap_y), bins = 200) + 
        labs(title = "Timing of monocyte record closest to serum collection", 
             # subtitle = "Time gap = 0 removed", 
             x = "Time gap in years", 
             caption = "n = 1049") + 
        theme_minimal() + 
        theme(axis.title = element_text(size = 12), 
              axis.text = element_text(size = 12), 
              plot.caption = element_text(size = 12))
ggplot(mono_closest) +
        geom_histogram(aes(mono), bins = 200) +
        labs(title = "Monocyte count closest to serum collection",
             x = "Mono (K/uL)",
             caption = "n = 1049") +
        theme_minimal() +
        theme(axis.title = element_text(size = 12),
              axis.text = element_text(size = 12),
              plot.caption = element_text(size = 12))

```

# 7. Save data

```{r}

# Save data table with aggregated info & one row per subject

wbc_count_lastbf <- merge(eos_lastbf[, .(EMPI, eos, eos_date, eos_unit, eos_timegap_d, 
                                         eos_timegap_y, eos_ref_range, eos_abnormal_flg)], 
                          neut_lastbf[, .(EMPI, neut, neut_date, neut_unit, neut_timegap_d, 
                                          neut_timegap_y, neut_ref_range, neut_abnormal_flg)], 
                          by = "EMPI", all = T)
wbc_count_lastbf <- merge(wbc_count_lastbf, 
                          baso_lastbf[, .(EMPI, baso, baso_date, baso_unit, baso_timegap_d, 
                                          baso_timegap_y, baso_ref_range, baso_abnormal_flg)], 
                          by = "EMPI", all = T)
wbc_count_lastbf <- merge(wbc_count_lastbf, 
                          lymph_lastbf[, .(EMPI, lymph, lymph_date, lymph_unit, lymph_timegap_d, 
                                           lymph_timegap_y, lymph_ref_range, lymph_abnormal_flg)], 
                          by = "EMPI", all = T)
wbc_count_lastbf <- merge(wbc_count_lastbf, 
                          mono_lastbf[, .(EMPI, mono, mono_date, mono_unit, mono_timegap_d, 
                                          mono_timegap_y, mono_ref_range, mono_abnormal_flg)], 
                          by = "EMPI", all = T)
fwrite(wbc_count_lastbf, file = here(dat_dir, "wbc_count_last_before_serum_collection.csv"))

wbc_count_closest <- merge(eos_closest[, .(EMPI, eos, eos_date, eos_unit, eos_timegap_d, 
                                           eos_timegap_y, eos_ref_range, eos_abnormal_flg)], 
                           neut_closest[, .(EMPI, neut, neut_date, neut_unit, neut_timegap_d, 
                                            neut_timegap_y, neut_ref_range, neut_abnormal_flg)], 
                           by = "EMPI", all = T)
wbc_count_closest <- merge(wbc_count_closest, 
                           baso_closest[, .(EMPI, baso, baso_date, baso_unit, baso_timegap_d, 
                                            baso_timegap_y, baso_ref_range, baso_abnormal_flg)], 
                           by = "EMPI", all = T)
wbc_count_closest <- merge(wbc_count_closest, 
                           lymph_closest[, .(EMPI, lymph, lymph_date, lymph_unit, lymph_timegap_d, 
                                             lymph_timegap_y, lymph_ref_range, lymph_abnormal_flg)], 
                           by = "EMPI", all = T)
wbc_count_closest <- merge(wbc_count_closest, 
                           mono_closest[, .(EMPI, mono, mono_date, mono_unit, mono_timegap_d, 
                                            mono_timegap_y, mono_ref_range, mono_abnormal_flg)], 
                           by = "EMPI", all = T)
fwrite(wbc_count_closest, file = here(dat_dir, "wbc_count_closest_to_serum_collection.csv"))


## Save long format data

fwrite(eos_long, file = here(long_fmt_dir, "eosinophil_long.csv"))
fwrite(neut_long, file = here(long_fmt_dir, "neutrophil_long.csv"))
fwrite(baso_long, file = here(long_fmt_dir, "basophil_long.csv"))
fwrite(lymph_long, file = here(long_fmt_dir, "lymphocyte_long.csv"))
fwrite(mono_long, file = here(long_fmt_dir, "monocyte_long.csv"))

```

# 8. Session info

```{r}

sessionInfo()

```
