---
title: "PEGASUS Analyses Accounting for Medications"
author: "MHuang"
date: "8/18/2020"
output: 
  html_document: 
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

## Packages

pkg <- c("tidyverse", "data.table", "here", "logging", "DT", "tableone", "survival", "gtools", 
         "ggrepel", "VennDiagram", "RColorBrewer")
for (p in pkg) {
        if (require(p, character.only = T)) {
                print(paste0(p, " loaded successfully"))
        } else {
                install.packages(p)
                require(p, character.only = T)
                print(paste0(p, " downloaded and loaded successfully"))
        }
}


## Paths

dat_dir <- "data/"
res_dir <- "results/"
fig_dir <- "figures/"

mike_dat_dir <- "C:/Users/MHuang/Dropbox (Partners HealthCare)/Asthma Data for Mike/06-30-2020/BiobankFiles/"


## Filenames

pegasus_dat_fname <- here(dat_dir, "updated_w_meds.RData")

biobank_file_name <- str_c(mike_dat_dir, "BiobankPortal_mns43_2020-06-30-123319.csv") 


## Hard-coded numbers

missing_thld <- 0.30
sig_digits <- 3
alpha_thld <- 0.05
z_95 <- qnorm(p = 0.975)

```

# 1. Check PPV

```{r}

process_deidentified <- function(input_file_name = biobank_file_name) {
        loginfo("Processing biobank file...")
        loginfo(str_c("file name: ", input_file_name))
        Deidentified <- fread(input_file_name)
        # Deidentified <- Deidentified %>% select(`Biobank Subject ID`, contains("asthma"))
        names(Deidentified) <- gsub("_$", "", 
                                    gsub(" |/", "_", 
                                         gsub("(\\(|\\)|-|Yes/No|Names/Text|All Concepts |List of All Values|\\[|\\]|\\?)", "", 
                                              gsub("Asthma - current or past history.+\\[>= 0.80PPV\\]", "Asthma_current_past_ppv80", 
                                                   names(Deidentified)))))
        Deidentified <- Deidentified %>%
                mutate_at(vars(contains("existence")), ~case_when(.x == "Yes" ~ 1, 
                                                                  .x == "No" ~ 0)) %>%
                mutate(Biobank_Subject_ID = as.integer(Biobank_Subject_ID)) %>% 
                as.data.table()
        loginfo(str_c(nrow(Deidentified), " subjects processed"))
        Deidentified[, asthma_ppv_num := gsub("\\[|\\]| ", "", Asthma_current_past_ppv80) %>% as.numeric()]
        if ("Plasma_Existence" %in% colnames(Deidentified)) {
                loginfo(str_c(nrow(Deidentified[Plasma_Existence == 1, ]), " subjects with plasma existence"))
        } else {loginfo("No info on plasma existence")}
        if ("Serum_Existence" %in% colnames(Deidentified)) {
                loginfo(str_c(nrow(Deidentified[Serum_Existence == 1, ]), " subjects with plasma existence"))
        } else {loginfo("No info on serum existence")}
        return(Deidentified)
}

all_asthma <- process_deidentified(biobank_file_name)

load(pegasus_dat_fname, verbose = T)

pegasus_meds[Group == "Asthmatic", Subject_Id] %in% all_asthma[, Biobank_Subject_ID] %>% table()
all_asthma[Biobank_Subject_ID %in% pegasus_meds[Group == "Asthmatic", Subject_Id], summary(asthma_ppv_num)]
all_asthma[Biobank_Subject_ID %in% pegasus_meds[Group == "Asthmatic", Subject_Id], .N, asthma_ppv_num >= 0.9]

```

# 2. metaboWAS - conditional logistic

```{r}

unmatched_stratum <- pegasus_meds[is.na(age), unique(stratum)]
pegasus_meds[stratum == unmatched_stratum, .(Subject_Id, stratum, Asthma_Status, Age, Gender, Race, BMI, 
                                             age, gender, race, bmi)] 
pegasus_meds[, .N, .(ocs_ard_serum_coll, cortisone_ard_serum_coll, Group)]
asth_to_rm <- pegasus_meds[ocs_ard_serum_coll == 1, Subject_Id]
pegasus_meds_trim <- pegasus_meds[stratum != unmatched_stratum & 
                                    !(Subject_Id %in% asth_to_rm) & !(Match_ID %in% asth_to_rm), ]

pegasus_noics_1y <- pegasus_meds_trim[stratum %in% pegasus_meds[Group == "Asthmatic" & ics_tot_1y == 0, stratum], ]
pegasus_ics_1y <- pegasus_meds_trim[stratum %in% pegasus_meds[Group == "Asthmatic" & ics_tot_1y >= 1, stratum], ]

pegasus_noocs_1y <- pegasus_meds_trim[stratum %in% pegasus_meds[Group == "Asthmatic" & ocs_tot_1y == 0, stratum], ]
pegasus_ocs_1y <- pegasus_meds_trim[stratum %in% pegasus_meds[Group == "Asthmatic" & ocs_tot_1y >= 1, stratum], ]


## Conditional logistic models ##

clogi_res_fnc <- function(outc, mets_list, dat, add_covar = "", 
                          first_cols = c("feature", "beta", "pval", "fdr_bh", "or", "lower95", "upper95", 
                                         "tentative_annotation", "pct_na_met", "cv", "totaln", "casen"), 
                          na_thld = missing_thld) {
        
        clogi_mdl_tmp <- function(met) {
                
                fit <- eval(parse(text = str_c("clogit(", outc, " ~ ", met, " + strata(stratum)", 
                                               add_covar, ", data = ", dat, ")")))
                
                coef <- summary(fit)$coefficients
                coef <- coef[rownames(coef) == met, ]
                
                confint <- summary(fit)$conf.int
                confint <- confint[rownames(confint) == met, -c(1:2)]
                
                totaln <- summary(fit)$n
                casen <- summary(fit)$nevent
                
                output <- t(c(outc, met, coef, confint, totaln, casen)) %>% as.data.table()
                colnames(output) <- c("outc", "feature", "beta", "or", "se", "zval", "pval", "lower95", "upper95", "totaln", "casen")
                output[, (colnames(output)[-c(1:2)]) := lapply(.SD, as.numeric), .SDcols = colnames(output)[-c(1:2)]]
                
                return(output)
        }
        
        res <- lapply(mets_list, function(x) {clogi_mdl_tmp(met = x)}) %>% rbindlist()
        res <- merge(res, 
                     mets_info[, .(feature, rt_min, mz, tentative_annotation, cv, pct_na_met, skewness_post)], 
                     by = "feature")
        res[pct_na_met <= na_thld, fdr_bh := p.adjust(pval, method = "BH")] # Only calculate FDR for features with <10% missing
        
        setcolorder(res, first_cols)
        res <- res[order(pval)]
        return(res)
}

feat_list <- mets_info[is.na(qc_rm) & pct_na_met <= missing_thld, feature] 
length(feat_list) 

res_crd <- clogi_res_fnc(outc = "asthma_sev", mets_list = feat_list, dat = "pegasus_meds_trim")


## ICS 

res_crd_noics_1y <- clogi_res_fnc(outc = "asthma_sev", mets_list = feat_list, dat = "pegasus_noics_1y")
ggplot(res_crd_noics_1y) + 
        geom_point(aes(x = beta, y = -log10(pval), color = tentative_annotation == "Unknown")) + 
        labs(title = "Asthmatics with no ICS within 1 year before blood collection vs matched controls (217 pairs)")
# ggplot(res_crd_noics_1y[tentative_annotation != "Unknown", ], aes(x = beta, y = -log10(pval))) + 
#         geom_point() + 
#         geom_label_repel(aes(label = ifelse(fdr_bh < alpha_thld, tentative_annotation, "")),
#                          box.padding   = 0.25, 
#                          point.padding = 0.5,
#                          segment.color = 'grey50')
# ggsave(here(fig_dir, "res_crd_noics_1y_annotated.png"), width = 12, height = 8)

res_crd_ics_1y <- clogi_res_fnc(outc = "asthma_sev", mets_list = feat_list, dat = "pegasus_ics_1y")
ggplot(res_crd_ics_1y) + 
        geom_point(aes(x = beta, y = -log10(pval), color = (tentative_annotation == "Unknown"))) + 
        labs(title = "Asthmatics with >=1 ICS within 1 year before blood collection vs matched controls (291 pairs)")
# ggplot(res_crd_ics_1y[tentative_annotation != "Unknown", ], aes(x = beta, y = -log10(pval))) + 
#         geom_point() + 
#         geom_label_repel(aes(label = ifelse(-log10(pval) > 4, tentative_annotation, "")),
#                          box.padding   = 0.25, 
#                          point.padding = 0.5,
#                          segment.color = 'grey50')
# ggsave(here(fig_dir, "res_crd_ics_1y_annotated.png"), width = 12, height = 8)

res_crd_noics_1y[pval < alpha_thld, feature] %in% res_crd_ics_1y[pval < alpha_thld, feature] %>% table()
res_crd_ics_1y[pval < alpha_thld, feature] %in% res_crd_noics_1y[pval < alpha_thld, feature] %>% table()

res_crd_noics_1y[fdr_bh < alpha_thld, feature] %in% res_crd_ics_1y[fdr_bh < alpha_thld, feature] %>% table()
res_crd_ics_1y[fdr_bh < alpha_thld, feature] %in% res_crd_noics_1y[fdr_bh < alpha_thld, feature] %>% table()

res_crd_noics_1y[pval < alpha_thld/length(feat_list), feature] %in% res_crd_ics_1y[pval < alpha_thld/length(feat_list), feature] %>% table()
res_crd_ics_1y[pval < alpha_thld/length(feat_list), feature] %in% res_crd_noics_1y[pval < alpha_thld/length(feat_list), feature] %>% table()


## OCS

res_crd_noocs_1y <- clogi_res_fnc(outc = "asthma_sev", mets_list = feat_list, dat = "pegasus_noocs_1y")
ggplot(res_crd_noocs_1y) + 
        geom_point(aes(x = beta, y = -log10(pval), color = tentative_annotation == "Unknown")) + 
        labs(title = "Asthmatocs with no OCS within 1 year before blood collection vs matched controls (217 pairs)")

res_crd_ocs_1y <- clogi_res_fnc(outc = "asthma_sev", mets_list = feat_list, dat = "pegasus_ocs_1y")
ggplot(res_crd_ocs_1y) + 
        geom_point(aes(x = beta, y = -log10(pval), color = (tentative_annotation == "Unknown"))) + 
        labs(title = "Asthmatocs with >=1 OCS within 1 year before blood collection vs matched controls (291 pairs)")

```

# 3. Within asthmatics - ICS/OCS & metabolites

```{r}

linear_res_fnc <- function(expo, mets_list, dat, add_covar = "", 
                           first_cols = c("feature", "beta", "pval", "fdr_bh", "tentative_annotation", 
                                          "pct_na_met", "cv"), 
                           na_thld = missing_thld) {
        
        linear_mdl_tmp <- function(met) {
                
                fit <- eval(parse(text = str_c("lm(", met, " ~ ", expo, " + Age + Gender + Race", 
                                               add_covar, ", data = ", dat, ")")))
                
                coef <- summary(fit)$coefficients
                coef <- coef[rownames(coef) == expo, ]

                # totaln <- summary(fit)$fstatistic[2] + summary(fit)$fstatistic[3] + 2
                
                output <- t(c(expo, met, coef)) %>% as.data.table()
                colnames(output) <- c("expo", "feature", "beta", "se", "tval", "pval")
                output[, (colnames(output)[-c(1:2)]) := lapply(.SD, as.numeric), .SDcols = colnames(output)[-c(1:2)]]
                
                return(output)
        }
        
        res <- lapply(mets_list, function(x) {linear_mdl_tmp(met = x)}) %>% rbindlist()
        res <- merge(res, 
                     mets_info[, .(feature, rt_min, mz, tentative_annotation, cv, pct_na_met, skewness_post)], 
                     by = "feature")
        res[pct_na_met <= na_thld, fdr_bh := p.adjust(pval, method = "BH")] # Only calculate FDR for metabolites with <10% missing
        
        setcolorder(res, first_cols)
        res <- res[order(pval)]
        return(res)
}

pegasus_meds_asth <- pegasus_meds_trim[Group == "Asthmatic", ]

res_mets_ics_1y <- linear_res_fnc(expo = "ics_tot_1y", mets_list = feat_list, dat = "pegasus_meds_asth")
res_mets_ics_2y <- linear_res_fnc(expo = "ics_tot_2y", mets_list = feat_list, dat = "pegasus_meds_asth")
res_mets_ics_5y <- linear_res_fnc(expo = "ics_tot_5y", mets_list = feat_list, dat = "pegasus_meds_asth")

res_mets_ocs_1y <- linear_res_fnc(expo = "ocs_tot_1y", mets_list = feat_list, dat = "pegasus_meds_asth")
res_mets_ocs_2y <- linear_res_fnc(expo = "ocs_tot_2y", mets_list = feat_list, dat = "pegasus_meds_asth")
res_mets_ocs_5y <- linear_res_fnc(expo = "ocs_tot_5y", mets_list = feat_list, dat = "pegasus_meds_asth")

ggplot(res_mets_ics_1y, aes(x = beta, y = -log10(pval))) + 
        geom_point(aes(color = tentative_annotation == "Unknown")) + 
        labs(title = "Metabolites association with ICS # of prescription within 1 year before blood collection (n=508)") + 
        geom_label_repel(aes(label = ifelse(tentative_annotation != "Unknown" & -log10(pval) > 2, tentative_annotation, "")),
                         box.padding   = 0.25, 
                         point.padding = 0.5,
                         segment.color = 'grey50')

res_crd[pval < alpha_thld, feature] %in% res_mets_ics_1y[pval < alpha_thld, feature] %>% table()
res_mets_ics_1y[pval < alpha_thld, feature] %in% res_crd[pval < alpha_thld, feature] %>% table()

res_crd[fdr_bh < alpha_thld, feature] %in% res_mets_ics_1y[fdr_bh < alpha_thld, feature] %>% table()
res_mets_ics_1y[fdr_bh < alpha_thld, feature] %in% res_crd[fdr_bh < alpha_thld, feature] %>% table()

res_crd[pval < alpha_thld/length(feat_list), feature] %in% res_mets_ics_1y[pval < alpha_thld/length(feat_list), feature] %>% table()
res_mets_ics_1y[pval < alpha_thld/length(feat_list), feature] %in% res_crd[pval < alpha_thld/length(feat_list), feature] %>% table()


## Venn diagrams

myCol <- brewer.pal(3, "Pastel2")

venn.diagram(list(res_crd_noics_1y[pval < alpha_thld, feature], 
                  res_crd_ics_1y[pval < alpha_thld, feature], 
                  res_mets_ics_1y[pval < alpha_thld, feature]), 
             category.names = c("Signif mets in non-ICS user" , "Signif mets in ICS user", 
                                "Mets assoc w/ # of ICS"),
             filename = "ics_1y_nominal_signif_venndiag.png",
             output=TRUE,
             
             # Output features
             imagetype="png" ,
             height = 1600 , 
             width = 1600 , 
             resolution = 300,
             compression = "lzw",
             
             # Circles
             lwd = 2,
             lty = 'blank',
             fill = myCol,
             
             # Numbers
             cex = 1,
             fontface = "bold",
             fontfamily = "sans",
             
             # Set names
             cat.cex = 1,
             cat.fontface = "bold",
             cat.default.pos = "outer",
             cat.pos = c(-18, 18, 135),
             cat.dist = c(0.055, 0.055, 0.085),
             cat.fontfamily = "sans",
             rotation = 1
)

venn.diagram(list(res_crd_noics_1y[fdr_bh < alpha_thld, feature], 
                  res_crd_ics_1y[fdr_bh < alpha_thld, feature], 
                  res_mets_ics_1y[fdr_bh < alpha_thld, feature]), 
             category.names = c("Signif mets in non-ICS user" , "Signif mets in ICS user", 
                                "Mets assoc w/ # of ICS"),
             filename = "ics_1y_fdr_signif_venndiag.png",
             output=TRUE,
             
             # Output features
             imagetype="png" ,
             height = 1600 , 
             width = 1600 , 
             resolution = 300,
             compression = "lzw",
             
             # Circles
             lwd = 2,
             lty = 'blank',
             fill = myCol,
             
             # Numbers
             cex = 1,
             fontface = "bold",
             fontfamily = "sans",
             
             # Set names
             cat.cex = 1,
             cat.fontface = "bold",
             cat.default.pos = "outer",
             cat.pos = c(-170, 170, 0),
             cat.dist = c(0.055, 0.055, 0.055),
             cat.fontfamily = "sans",
             rotation = 1
)

venn.diagram(list(res_crd_noics_1y[pval < alpha_thld/length(feat_list), feature], 
                  res_crd_ics_1y[pval < alpha_thld/length(feat_list), feature], 
                  res_mets_ics_1y[pval < alpha_thld/length(feat_list), feature]), 
             category.names = c("Signif mets in non-ICS user" , "Signif mets in ICS user", 
                                "Mets assoc w/ # of ICS"),
             filename = "ics_1y_bonf_signif_venndiag.png",
             output=TRUE,
             
             # Output features
             imagetype="png" ,
             height = 1600 , 
             width = 1600 , 
             resolution = 300,
             compression = "lzw",
             
             # Circles
             lwd = 2,
             lty = 'blank',
             fill = myCol,
             
             # Numbers
             cex = 1,
             fontface = "bold",
             fontfamily = "sans",
             
             # Set names
             cat.cex = 1,
             cat.fontface = "bold",
             cat.default.pos = "outer",
             cat.pos = c(-180, 0, 180),
             cat.dist = c(0.055, 0.055, 0.085),
             cat.fontfamily = "sans",
             rotation = 1
)

# venn.diagram(list(res_crd_noocs_1y[pval < alpha_thld, feature], 
#                   res_crd_ocs_1y[pval < alpha_thld, feature], 
#                   res_mets_ocs_1y[pval < alpha_thld, feature]), 
#              category.names = c("Signif mets in non-OCS user" , "Signif mets in OCS user", 
#                                 "Mets assoc w/ # of OCS"),
#              filename = "ocs_1y_nominal_signif_venndiag.png",
#              output=TRUE,
#              
#              # Output features
#              imagetype="png" ,
#              height = 1600 , 
#              width = 1600 , 
#              resolution = 300,
#              compression = "lzw",
#              
#              # Circles
#              lwd = 2,
#              lty = 'blank',
#              fill = myCol,
#              
#              # Numbers
#              cex = 1,
#              fontface = "bold",
#              fontfamily = "sans",
#              
#              # Set names
#              cat.cex = 1,
#              cat.fontface = "bold",
#              cat.default.pos = "outer",
#              cat.pos = c(-18, 18, 135),
#              cat.dist = c(0.055, 0.055, 0.085),
#              cat.fontfamily = "sans",
#              rotation = 1
# )
# 
# venn.diagram(list(res_crd_noocs_1y[fdr_bh < alpha_thld, feature], 
#                   res_crd_ocs_1y[fdr_bh < alpha_thld, feature], 
#                   res_mets_ocs_1y[fdr_bh < alpha_thld, feature]), 
#              category.names = c("Signif mets in non-OCS user" , "Signif mets in OCS user", 
#                                 "Mets assoc w/ # of OCS"),
#              filename = "ocs_1y_fdr_signif_venndiag.png",
#              output=TRUE,
#              
#              # Output features
#              imagetype="png" ,
#              height = 1600 , 
#              width = 1600 , 
#              resolution = 300,
#              compression = "lzw",
#              
#              # Circles
#              lwd = 2,
#              lty = 'blank',
#              fill = myCol,
#              
#              # Numbers
#              cex = 1,
#              fontface = "bold",
#              fontfamily = "sans",
#              
#              # Set names
#              cat.cex = 1,
#              cat.fontface = "bold",
#              cat.default.pos = "outer",
#              cat.pos = c(-170, 170, 0),
#              cat.dist = c(0.055, 0.055, 0.055),
#              cat.fontfamily = "sans",
#              rotation = 1
# )
# 
# venn.diagram(list(res_crd_noocs_1y[pval < alpha_thld/length(feat_list), feature], 
#                   res_crd_ocs_1y[pval < alpha_thld/length(feat_list), feature], 
#                   res_mets_ocs_1y[pval < alpha_thld/length(feat_list), feature]), 
#              category.names = c("Signif mets in non-OCS user" , "Signif mets in OCS user", 
#                                 "Mets assoc w/ # of OCS"),
#              filename = "ocs_1y_bonf_signif_venndiag.png",
#              output=TRUE,
#              
#              # Output features
#              imagetype="png" ,
#              height = 1600 , 
#              width = 1600 , 
#              resolution = 300,
#              compression = "lzw",
#              
#              # Circles
#              lwd = 2,
#              lty = 'blank',
#              fill = myCol,
#              
#              # Numbers
#              cex = 1,
#              fontface = "bold",
#              fontfamily = "sans",
#              
#              # Set names
#              cat.cex = 1,
#              cat.fontface = "bold",
#              cat.default.pos = "outer",
#              cat.pos = c(-180, 0, 180),
#              cat.dist = c(0.055, 0.055, 0.085),
#              cat.fontfamily = "sans",
#              rotation = 1
# )

```

# 4. Save results

```{r}

save(list = c(grep("^res_", ls(), value = T), grep("^pegasus_", ls(), value = T), "mets_info", "feat_list"), 
     file = here(res_dir, "results.RData"))

```

## 4.1. Overall conditional logistic results

```{r}

datatable(res_crd[, c(1:13)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)
datatable(res_crd[tentative_annotation != "Unknown", c(1:13)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)

```

## 4.2. Asthmatics with no ICS within 1 year before blood collection vs matched controls (217 pairs)

```{r}

datatable(res_crd_noics_1y[, c(1:13)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)
datatable(res_crd_noics_1y[tentative_annotation != "Unknown", c(1:13)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)

```

## 4.3. Asthmatics with >=1 ICS within 1 year before blood collection vs matched controls (291 pairs)

```{r}

datatable(res_crd_ics_1y[, c(1:13)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)
datatable(res_crd_ics_1y[tentative_annotation != "Unknown", c(1:13)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)

```

## 4.3. Asthmatics with >=1 ICS within 1 year before blood collection vs matched controls (291 pairs)

```{r}

datatable(res_crd_ics_1y[, c(1:13)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)
datatable(res_crd_ics_1y[tentative_annotation != "Unknown", c(1:13)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)

```

## 4.4. Metabolites association with # of ICS prescription within 1 year before blood collection (n=508)

```{r}

datatable(res_mets_ics_1y[, c(1:8)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "pct_na_met", "cv"), digits = 3)
datatable(res_mets_ics_1y[tentative_annotation != "Unknown", c(1:8)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "pct_na_met", "cv"), digits = 3)

```

## 4.5. Metabolites association with # of ICS prescription within 1 year before blood collection (n=508)

```{r}

datatable(res_mets_ics_1y[, c(1:8)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "pct_na_met", "cv"), digits = 3)
datatable(res_mets_ics_1y[tentative_annotation != "Unknown", c(1:8)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "pct_na_met", "cv"), digits = 3)

```

## 4.6. Metabolites association with # of ICS prescription within 2 years before blood collection (n=508)

```{r}

datatable(res_mets_ics_2y[, c(1:8)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "pct_na_met", "cv"), digits = 3)
datatable(res_mets_ics_2y[tentative_annotation != "Unknown", c(1:8)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "pct_na_met", "cv"), digits = 3)

```

## 4.7. Metabolites association with # of ICS prescription within 5 years before blood collection (n=508)

```{r}

datatable(res_mets_ics_5y[, c(1:8)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "pct_na_met", "cv"), digits = 3)
datatable(res_mets_ics_5y[tentative_annotation != "Unknown", c(1:8)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "pct_na_met", "cv"), digits = 3)

```

# 5. Session info

```{r}

sessionInfo()

```
